Predicting Transport in Lyotropic Liquid Crystal Membranes with Molecular Dynamics Simulations -- Outline

### 2016-12-19 Ben_Manuscripts/Paper1/Manuscript.tex
Finished text for methods section. Started text for introduction

\documentclass{article}
\title{Predicting Transport in Lyotropic Liquid Crystal Membranes with Molecular Dynamics Simulations -- Outline}
\author{Benjamin J. Coscia \and Douglas L. Gin \and Richard D. Noble \and Michael R. Shirts} 
\begin{document}
	\bibliographystyle{ieeetr}
	\maketitle
	\section{Introduction}
	Nanostructured membrane materials have become increasingly popular in desalination and wastewater treatment applications because they offer the ability to control membrane architecture at the atomic scale.
	\begin{itemize}
		\item Current state-of-the-art reverse osmosis membranes are dense and unstructured with tortuous and polydisperse pores which lead to inconsistent performance
		\item Tortuosity and polydispersity drive up energy requirements which strain developing regions and contribute strongly to CO2 emissions
		\item With nanostructured materials, solute rejecting pores can be tuned uniformly -- drives down energy requirements
	\end{itemize}

	Development of nanostructured materials has been limited by the ability to synthesize and scale various fundamentally sound technologies.
	\begin{itemize}
		\item Leading technologies and their limitations:
		\begin{itemize}
			\item Graphene sheets -atomically thick which gives excellent permeability but defects during manufacturing severely impact performance
			\item Carbon Nanotubes - MD studies are promising but synthetic techniques unable to achieve necessary alignment and pore monodispersity
		\end{itemize}
	\end{itemize} 
	Self assembling lyotropic liquid crystals (LLCs) share the characteristic ability of nanostructured membrane materials to create highly ordered structures with the benefits of low cost and synthetic techniques feasible for large scale production.
	\begin{itemize}
		\item Forms lamellar, bicontinuous cubic and hexagonal phase based on solution composition
		\item My monomer form two types of self assembled LLCs - thermotropic (Colh) and lyotropic (HII) 
		\item The thermotropic, Colh is formed by the self assembly of neat monomer
		\item The lyotropic, HII phase is formed in the presence of small percentages of water (around 8 percent)
		\item Both assemble into cyclinders with hydrophilic groups oriented inward towards the pore center and hydrophobic group facing outward. The only difference is the inclusion of water in the structure which leads to minor variations in the structure with potentially different filtration properties (although no filtration experiments have been done on Colh)  
		\item Hydrophilic regions point towards pore centers
		\item Until recently, they could not be aligned - hindered progress
		\item Yale aligns them, then crosslinks them to lock in the structure
		\item LLC HII phase membranes offer potential for high permeability and selectivity which equals low energy consumption
		\item This paper will focus on the development of a model of the colh phase since it is a simpler starting point and has just as much experimental data
	\end{itemize}
	Constructing a molecular model may give a clearer picture of the microscopic structure of LLC membranes which will change how we think about their design.
	\begin{itemize}
		\item The arrangement of sodium ions in the channels is thought to be confined to the pore walls. It is possible they are arranged more randomly
		\begin{itemize}
			\item This could change how one thinks about molecules diffusing through membrane
			\item Could also be a difference between lyotropic and thermotropic phases
		\end{itemize}
		\item How many monomers are in a layer?
		\begin{itemize}
			\item There is no clear answer to this question in literature. I'd like to confirm or deny past predictions
			% BJC: It seems that the following discussion should be placed in the results section, with the above lines belonging in this section, posing the question along with past work/justification
			\item It has been suggested that there are 4 monomers in a each disk/layer, however, simulations have shown that this leads to unstable configurations with dimensions that are too small compared to experiment.
			\item Stable simulations with 6 and 7 monomers have been performed (will do one with five soon) which give structural characteristics consistent with experiment
			\item Stable systems have also been simulated consisting of varying numbers of monomers per layer
			\item This suggests that the arrangement of monomers is more complicated than simple layers stacked on top of each other
		\end{itemize}
	\end{itemize}

@@

	The timescale for self assembly of monomers into the hexagonal phase is unknown and likely outside of a reasonable simulation length. Past work has shown the ability of other liquid crystals to form into various phases including the hexagonal phase using coarse grain models in c.a. 1000 ns. While 1000 ns would be expensive with an atomistic model, it is not an unreasonable amount of time and prompted an attempt at a self assembly. 480 monomers were packed into a cubic box using Packmol software, with box dimensions chosen based on approximate density of the col\subscript{h} phase. The box was made slightly larger than this to help with Packmol convergence and to give the monomers room to reorient themselves during equilibration. NPT simulations were carried out for 100 ns. In that time, no visible progress was made towards self assembly so simulations were stopped. Even if self assembly happened eventually, it is not a sustainable way to set up the system due to computational cost.

@@

        The validity of our model can be tested based on its agreement with experimental structural and material property data. Our simulations should mimic the real system as best as possible within the limits imposed by choice of force field. This system is particularly difficult since there has not been a forcefield developed for it and the system has not been studied in the past. Therefor, parameters were chosen carefully based on the ability to reproducibly create acceptably accurate structures.

@@

	The spacing between layers has a significant impact on the system dimensions. It should be between 4 A and 7 A, with a recommended value of 5 A. A spacing below 4 angstroms risks large repulsions between layers during the first steps of simulation giving unpredictable results. A spacing greater than 7 A has consistently given structural dimensions that are too small. It is hypothesized that it is due to monomer layers slipping between each during the first picoseconds of simulation. The system gets locked into a metastable state from which there is no recovery. A spacing of 5 A allows a more gentle relaxation and subsequent relaxation into a experimentally consistent configuration.

	Experimental small angle X-ray scattering results were used to inform our choice of an initial pore to pore distance. The peak corresponding to the d100 plane was used to estimate a distance between pores of 4.12 nm. The sensitivity of the system to the starting pore-to-pore distance was tested by simulating initial configurations with starting pore-to-pore distances of 3.0, 3.5, 4.0 and 4.5 nm. Using 3.0 and 3.5 nm resulted in a structure whose dimensions were too small. The significant overlap between pores at these initial conditions likely locked the pores together with no hope of separation without energy input. Starting pore-to-pore distances of 4.0 and 4.5 nm both converged to expected distances.   
	\begin{itemize}
		\item SAXS to get an idea of Pore-to-Pore
		\begin{itemize}
			\item multiple starting pore to pore distances tested (3.0, 3.5, 4.0, 4.5 nm)
			\item 4.0 nm and 4.5 nm converged to similar dimensions, 4.5 nm was most reliable
		\end{itemize}
		\item TEM images and rejection give a pore size estimate
	\end{itemize}

	\noindent NPT simulations were run at 300 K and 1 bar for system equilibration.
	\begin{itemize}
		\item semi-isotropic
		\item Monitored 'hexagonality' - i.e. how the distance between pores compares to an ideal hexagon
		\item Varied initial pore-to-pore distance and ran very long simulations - typically > 200 ns
	\end{itemize}

	After an equilibrated structure was created, a crosslinking procedure was performed in order to better parallel synthetic procedures. 
	\begin{itemize}
		\item Crosslinking maintains alignment of cylindrical mesophases
		\item head to tail addition
		\item racemic mixture 
		\item Details of crosslinking algorithm (refer to appendix or supplemental info)
	\end{itemize}

>>>>

\documentclass{article}
\title{Predicting Transport in Lyotropic Liquid Crystal Membranes with Molecular Dynamics Simulations -- Outline}
\author{Benjamin J. Coscia \and Douglas L. Gin \and Richard D. Noble \and Michael R. Shirts} 
\begin{document}
	\bibliographystyle{ieeetr}
	\maketitle
	\section{Introduction}
	Nanostructured membrane materials have become increasingly popular in desalination and wastewater treatment applications because they offer the ability to control membrane architecture at the atomic scale. Current state-of-the-art reverse osmosis membranes are dense and unstructured with tortuous and polydisperse pores. Polydispersity in pore radius limits the ability to completely reject solutes of a given size. Pore tortuosity decreases membrane permeability requiring high feed pressures in order to reach a reasonable throughput. Nanostructured materials can circumvent both issues. Theoretically one can precisely design solute rejecting pores that are uniform in size and run straight through the membrane resulting in optimal rejection and permeability.

	Development of nanostructured materials has been limited by the ability to synthesize and scale various fundamentally sound technologies. Two notable examples are graphene sheets and carbon nanotubes. Graphene sheets are atomically thick and pores can be made uniform in size, however defects during manufacturing severly impact solute rejection. Currently, some researchers are investigating multilayered graphene sheets which should be more mechanically robust and still provide high performance. Carbon nanotubes have shown great promise from molecular dynamics simulations exhibiting surprisingly high water permeability, but current synthetic techniques are unable to achieve the necessary alignment and pore monodispersity which would set them apart from other technologies.

	Self assembling lyotropic liquid crystals (LLCs) share the characteristic ability of nanostructured membrane materials to create highly ordered structures with the benefits of low cost and synthetic techniques feasible for large scale production.
	\begin{itemize}
		\item Forms lamellar, bicontinuous cubic and hexagonal phase based on solution composition
		\item My monomer form two types of self assembled LLCs - thermotropic (Colh) and lyotropic (HII) 
		\item The thermotropic, Colh is formed by the self assembly of neat monomer
		\item The lyotropic, HII phase is formed in the presence of small percentages of water (around 8 percent)
		\item Both assemble into cyclinders with hydrophilic groups oriented inward towards the pore center and hydrophobic group facing outward. The only difference is the inclusion of water in the structure which leads to minor variations in the structure with potentially different filtration properties (although no filtration experiments have been done on Colh)  
		\item Hydrophilic regions point towards pore centers
		\item Until recently, they could not be aligned - hindered progress
		\item Yale aligns them, then crosslinks them to lock in the structure
		\item LLC HII phase membranes offer potential for high permeability and selectivity which equals low energy consumption
		\item This paper will focus on the development of a model of the colh phase since it is a simpler starting point and has just as much experimental data
	\end{itemize}
	Constructing a molecular model may give a clearer picture of the microscopic structure of LLC membranes which will change how we think about their design.
	\begin{itemize}
		\item The arrangement of sodium ions in the channels is thought to be confined to the pore walls. It is possible they are arranged more randomly
		\begin{itemize}
			\item This could change how one thinks about molecules diffusing through membrane
			\item Could also be a difference between lyotropic and thermotropic phases
		\end{itemize}
		\item How many monomers are in a layer?
		\begin{itemize}
			\item There is no clear answer to this question in literature. I'd like to confirm or deny past predictions
			% BJC: It seems that the following discussion should be placed in the results section, with the above lines belonging in this section, posing the question along with past work/justification
			\item It has been suggested that there are 4 monomers in a each disk/layer, however, simulations have shown that this leads to unstable configurations with dimensions that are too small compared to experiment.
			\item Stable simulations with 6 and 7 monomers have been performed (will do one with five soon) which give structural characteristics consistent with experiment
			\item Stable systems have also been simulated consisting of varying numbers of monomers per layer
			\item This suggests that the arrangement of monomers is more complicated than simple layers stacked on top of each other
		\end{itemize}
	\end{itemize}

@@

	The timescale for self assembly of monomers into the hexagonal phase is unknown and likely outside of a reasonable simulation length. Past work has shown the ability of other liquid crystals to form into various phases including the hexagonal phase using coarse grain models in c.a. 1000 ns. While 1000 ns would be expensive with an atomistic model, it is not an unreasonable amount of time and prompted an attempt at a self assembly. 480 monomers were packed into a cubic box using Packmol software, with box dimensions chosen based on approximate density of the col\subscript{h} phase. The box was made about 10 percent larger than this to help with Packmol convergence and to give the monomers room to reorient themselves during equilibration. NPT simulations were carried out for 100 ns. In that time, no visible progress was made towards self assembly so simulations were stopped. Even if self assembly happened eventually, it is not a computationally sustainable way to set up the system.

@@

        The validity of our model can be tested based on its agreement with experimental structural and material property data. Our simulations should mimic the real system as best as possible within the limits imposed by choice of force field. This system is particularly difficult since there has not been a forcefield developed for it and the system has not been studied in the past. Therefore, parameters were chosen carefully based on the ability to reproducibly create acceptably accurate structures.

@@

	The spacing between layers has a significant impact on the system dimensions. It should be between 4 A and 7 A, with a recommended value of 5 A. A spacing below 4 angstroms risks large repulsions between layers during the first steps of simulation giving unpredictable results. A spacing greater than 7 A has consistently given structural dimensions that are too small. It is hypothesized that it is due to monomer layers slipping between each other during the first picoseconds of simulation. The system gets locked into a metastable state from which there is no recovery. A spacing of 5 A allows a more gentle equilibration and subsequent relaxation into a experimentally consistent configuration.

	Experimental small angle X-ray scattering results were used to inform our choice of an initial pore to pore distance. The peak corresponding to the d\textsubscript{100} plane was used to estimate a distance between pores of 4.12 nm. The sensitivity of the system to the starting pore-to-pore distance was tested by simulating initial configurations with starting pore-to-pore distances of 3.0, 3.5, 4.0 and 4.5 nm. Using 3.0 and 3.5 nm resulted in a structure whose dimensions were too small. The significant overlap between pores at these initial conditions likely locked the pores together with no hope of separation without energy input. Starting pore-to-pore distances of 4.0 and 4.5 nm both converged to expected distances.

	Experimental pore radius measurements do not currently give a clear connection to simulations. Rejection studies on the H\subscript{II} phase showed the ability of the pores to completely reject solutes 1.2 nm in diameter. TEM images confirmed the result by approximating a pore radius based on the fluorescence of stained benzene rings in the LLC monomer. For that reason, the initial pore radius was initialized at 0.6 nm. A clear definition of the pore boundaries is yet to be defined experimentally and measurements based on simulations are based on the author's methodology until a consistent method is agreed upon.

	NPT simulations run at 300 K and 1 bar with semistropic box vectors and periodic boundaries on in three dimensions. Equilibration was monitored quantatively by observing the distance between pores over time. Once the distances stopped changing (disregarding equilibrium fluctuations) the system was determined to be equilibrated. Typical equilibrations take about 200 ns but are run for up to 500 ns or longer to ensure our results.

	After an equilibrated structure was created, a crosslinking procedure was performed in order to better parallel synthetic procedures. The primary function of crosslinking in the col\textsubscript{h} phase is to maintain its ordered structure and mechanical stability. A simple crosslinking algorithm was implemented to created a structure which mimics the real system. All crosslinking occurrs at the vinyl groups present at the end of each monomer tail. Only head-to-tail addition was considered since it is the mode of addition which dominates. Because the reaction is free radical polymerization, atoms are necessarily added to the system. This was achieved using dummy atoms meant to represent initiator and added hydrogens. The initiator was simulated as hydrogen for simplicity and should have no effect on our results because of its dilute presence. The overall procedure was repeated iteratively until the desired crosslinking density of 95 \% was achieved. During each iteration, eligible bonding carbons were determined based on their distance from other potentially bonding carbons. A specified percent of the distribution of eligible carbons are bonded and the topology is updated accordingly. Since the resultant crosslinked mixture is racemic and we are not interested in the kinetics, the orientation of bonding carbons was not considered. Following each iteration, a short simulation of 50 ps is run to reposition all atoms. Please refer to the supplemental information for detail about the algorithm.
